---
title: "Merge, aggregate and count in R"
---
  
The purpose of this section is to show some examples of SQL like data analysis using R. The data that is used in these examples is the Customer Support data from [IBM Watson Analytics](https://www.ibm.com/communities/analytics/watson-analytics-blog/guide-to-sample-datasets/).   
  
First, we want to load the libraries that we need:
```{r}
library(data.table)
library(plyr)
```


```{r}
churnAnalysis <- fread("WA_Fn-UseC_-Telco-Customer-Churn.csv") 
churnAnalysis <- as.data.frame(churnAnalysis)
head(churnAnalysis)
```

### Aggregate function 

Let's say, if we are interested in if payment method is a driver for monthly charges, we can use **aggregate** function to see the average monthly charges of each payment method:
```{r, message=F, warning=F}
PaymentMethodMonthlyChargesMean <- aggregate(churnAnalysis[,c("PaymentMethod","MonthlyCharges")], list(churnAnalysis$PaymentMethod), mean)
PaymentMethodMonthlyChargesMean
```
People who use mailed check to pay seems to have lower monthly charges.

### Count function

From the data, we observe that the people do mailed check has lower monthly charges, but to make sure that it is not driven by a small number of data, we do **count** for count how many data points we have for each payment methods:
```{r}
count(churnAnalysis,'PaymentMethod') ## library plyr
```
Given that each payment mehtod has more than 1000 data points, we can say that the difference is not driven by a small number of data.

```{r}
library("RSQLite")
# connect to the sqlite file
mydb = dbConnect(RSQLite::SQLite(), dbname="database.sqlite") ## load the music review data
## list all tables
tableList <- dbListTables(mydb)



# dbDisconnect(mydb)
# unlink("my-db.sqlite")
```

```{r}
reviews <- dbReadTable(mydb,"reviews")
artists <- dbReadTable(mydb,"artists")
## I want to find out how many artists first
artistsList <-  dbGetQuery(mydb, 'SELECT DISTINCT artist  FROM artists')
## I want to find out for each artist, how many reviews have they got (use Group By)
artistsReviewCount <- dbGetQuery(mydb, 'SELECT COUNT(a.reviewid), a.* FROM artists AS a GROUP BY a.artist ORDER BY COUNT(a.reviewid) DESC')
head(artistsReviewCount)
```



```{r}
##use R to solve the SQL problem
## I want to find out how many artists first
artistList <- unique(artists)
#tempReview_Artist <- merge(artists,reviews,by.x="reviewid",by.y="reviewid",all.y=TRUE) ## artist is already in the review table....
## I want to find out for each artist, how many reviews have they got
reviewCounts <- count(artists,"artist")
## only want to see artists with more than 1 reviews
reviewCountsEx1 <- reviewCounts[reviewCounts[,2]>1,]
```

```{r}
## let's say, I don't have plyr library, what can I do?
countGroupBy <- function(table,ByColumn) {
  groups <- unique(table[,ByColumn])
  #NewTable <- as.data.frame(matrix(NA,nrow=length(groups),ncol=2))
  temp <- sapply(groups,function(x,table,ByColumn) {return(length(which(table[,ByColumn] ==x)))},table=table,ByColumn=ByColumn)
  return(cbind(groups,as.numeric(temp)))
}
reviewCountsBySelf <- countGroupBy(table=artists,ByColumn="artist")
reviewCountsBySelf <- reviewCountsBySelf[order(reviewCountsBySelf[,1]),]
reviewCounts <- reviewCounts[order(reviewCounts[,1]),]
#sum(reviewCounts!=reviewCountsBySelf) ## check if the self-built function is the same as
## the count function from plyr package
```

```{r}
## use SQL
## I want to find out for each artist, what's the album (order) that has the highest score
years <- dbGetQuery(mydb, 'SELECT *  FROM years')
artistsList <-  dbGetQuery(mydb, 'SELECT DISTINCT artist  FROM artists')
dbSendQuery(mydb, 'CREATE TABLE releaseYears AS SELECT reviewid,MIN(year) AS releaseYear FROM years GROUP BY years.reviewid')
releaseYears <- dbGetQuery(mydb, 'SELECT * FROM releaseYears')
dbGetQuery(mydb, 'SELECT * FROM releaseYears WHERE releaseYear IS NULL') 
dbSendQuery(mydb, 'DELETE FROM releaseYears WHERE releaseYear IS NULL') ## remove the review that doesn't have release year


## I want to find out for each artist, how many reviews have they got (use Group By)
# artistsReviewScore <- dbGetQuery(mydb, 'SELECT ROW_NUMBER() OVER(PARTITION BY[r.artist] ORDER BY r.pub_year) FROM reviews AS r') ## ROW_NUMBER() doesn't work for RSQLite
dbSendQuery(mydb, 'CREATE TABLE newreviews AS SELECT r.*,(SELECT count(*) FROM (SELECT reviews.*,releaseYears.releaseYear FROM reviews JOIN releaseYears on reviews.reviewid=releaseYears.reviewid) AS r2 WHERE r.artist=r2.artist AND r.releaseYear >= r2.releaseYear) AS orderIndex FROM (SELECT reviews.*,releaseYears.releaseYear FROM reviews JOIN releaseYears ON reviews.reviewid=releaseYears.reviewid) AS r')


artistsReviewOrder <- dbGetQuery(mydb, 'SELECT * FROM newreviews')
## test using massive attack
artistsReviewOrder[which(artistsReviewOrder[,"artist"]=="massive attack"),]
## test using oasis
artistsReviewOrder[which(artistsReviewOrder[,"artist"]=="oasis"),]
## test using muse
artistsReviewOrder[which(artistsReviewOrder[,"artist"]=="muse"),]

artistsHighestReviewAlbum <- dbGetQuery(mydb, 'SELECT max(t.score) AS score,count(*) AS numAlbums,t.artist,t.title,t.orderIndex,t.releaseYear FROM newreviews t GROUP BY t.artist')

## use massive attack to test
artistsHighestReviewAlbum[which(artistsHighestReviewAlbum[,"artist"] %in% c("massive attack","radiohead","arctic monkeys","muse","oasis")),c("score","title","artist","orderIndex","releaseYear","numAlbums")]

## plot the hist for artists that have >= 5 albums
hist(artistsHighestReviewAlbum[which(artistsHighestReviewAlbum[,"numAlbums"]>=5),"orderIndex"]-0.01,xlim = c(0,15),breaks=100)
```

```{r}
genres <- dbGetQuery(mydb, 'SELECT g.* FROM genres AS g')
reviewWithGenres <- dbGetQuery(mydb, 'SELECT r.reviewid, r.title, r.artist, r.score, r.releaseYear, r.orderIndex, g.genre FROM newreviews AS r JOIN genres AS g on g.reviewid=r.reviewid')

dbSendQuery(mydb, 'CREATE TABLE reviewWithGenres AS SELECT r.reviewid, r.title, r.artist,r.score, r.releaseYear, r.orderIndex, g.genre FROM newreviews AS r JOIN genres AS g on g.reviewid=r.reviewid')
reviewWithGenres <- dbGetQuery(mydb, 'SELECT * FROM reviewWithGenres')
GenresAvgScore <- dbGetQuery(mydb, 'SELECT AVG(r.score),r.genre,count(*) FROM reviewWithGenres r GROUP BY r.genre')
## find the ratio of scores >= 7 for each group
dbGetQuery(mydb, 'WITH temp AS (SELECT r.genre, count(*) AS countHigh FROM reviewWithGenres r WHERE r.score >= 7 GROUP BY r.genre), 
           temp2 AS (SELECT r. genre, count(*) AS countAll FROM reviewWithGenres r GROUP BY r.genre)
           SELECT a.genre, a.countAll, h.countHigh, CAST(h.countHigh AS DOUBLE)/a.countAll FROM temp h JOIN temp2 a ON a.genre = h.genre')


## how many review doesn't have genre (check NULL)
dbGetQuery(mydb,'SELECT count(*) FROM genres g WHERE g.genre IS NULL')
dbGetQuery(mydb,'SELECT count(*) FROM reviewWithGenres g WHERE g.genre IS NULL')
## where is the 67 rows in genre table but not in reviewWithGenres table?

## important
dbGetQuery(mydb, 'WITH allCount AS (SELECT r.genre, count(*) AS countA FROM reviewWithGenres r GROUP BY r.genre)
, highScoreCount AS (SELECT r.genre, count(*) AS countH FROM reviewWithGenres r WHERE r.score >= 7 GROUP BY r.genre)
SELECT a.genre, CAST(h.countH AS DOUBLE)/a.countA ,h.countH, a.countA FROM allCount a JOIN highScoreCount h ON a.genre=h.genre OR (a.genre IS NULL AND h.genre IS NULL)')



## find the question with highest conversion ratio
## SELECT Qid, CAST(count(CASE WHEN event=answered THEN 1) AS DOUBLE)/count(CASE WHEN event = saw THEN 1) FROM table t GROUP BY t.Qid

## WITH temp AS (SELECT userName, Qid, event, order FROM table t WHERE Qid = i AND event = answer), 
## tempNextShown AS (SELECT Qid,count(*) AS ShownCount FROM table t JOIN temp ON  t.order = temp.order + 1 AND event = "saw" AND t.userName = temp.userName GROUP BY Qid),
## tempNextAnswer AS (SELECT Qid,count(*) AS AnswerCount FROM table t JOIN temp ON t.order = temp.order + 1 AND event = "answered" AND t.userName = temp.userName GROUP BY Qid) 
## SELECT tempNextShown.Qid, CAST(AnswerCount AS DOUBLE)/ShownCount, FROM tempNextShown JOIN tempNextAnswer ON tempNextShown.Qid = tempNextAnswer.Qid

## No no no, forget about the order, just find the user that skip question i, and for those users, find the question that has highest conversion ratio
## WITH temp AS (SELECT userName, Qid, event FROM table t WHERE Qid = i AND event = answered), 
## Saw AS (SELECT Qid, count(*) AS countSaw FROM table t JOIN temp ON t.userName = temp.userName WHERE event = saw GROUP BY Qid), 
## A AS (SELECT Qid, count(*) AS countA FROM table t JOIN temp ON t.userName = temp.userName WHERE event = answerd GROUP BY Qid)
## SELECT Qid, CAST(countA AS DOUBLE)/countSaw FROM Saw JOIN A ON Saw.Qid = A.Qid




## For the artists that their first album's score is higher than 6, what's the probability that next album's score is also higher than 6

dbGetQuery(mydb,'WITH temp AS (SELECT g.reviewid, g.artist, g.title FROM reviewWithGenres g WHERE g.artist IN ("muse","radiohead"))
           SELECT r.reviewid, r.title, r.artist FROM temp, reviewWithGenres r WHERE r.reviewid = temp.reviewid')
dbGetQuery(mydb,'WITH temp AS (SELECT g.reviewid, g.artist, g.title FROM reviewWithGenres g WHERE g.artist IN ("muse","radiohead"))
           SELECT r.reviewid, r.title, r.artist FROM temp, reviewWithGenres AS r WHERE r.artist IN (SELECT DISTINCT artist FROM temp)')
## when table temp is not defined in WHERE, then database is not working as expected
dbGetQuery(mydb,'WITH temp AS (SELECT g.reviewid, g.artist, g.title FROM reviewWithGenres g WHERE g.artist IN ("muse","radiohead"))
           SELECT r.reviewid, r.title, r.artist FROM reviewWithGenres AS r WHERE r.artist IN (SELECT DISTINCT artist FROM temp)')
## now it's better

## given that the highest score, what's the probability that next album's score is >=7
dbGetQuery(mydb,'WITH temp AS (SELECT r.reviewid, r.title, r.artist, r.genre,  MAX(r.score), r.orderIndex FROM reviewWithGenres r GROUP BY r.artist), 
           temp2 AS (SELECT r.genre, count(*) AS countAll FROM reviewWithGenres r JOIN temp ON r.orderIndex = temp.orderIndex+1 AND r.artist = temp.artist GROUP BY r.genre),
           temp3 AS (SELECT r.genre, count(*) AS countHigh FROM reviewWithGenres r JOIN temp ON r.orderIndex = temp.orderIndex+1 AND r.artist = temp.artist AND r.score >=7.5 GROUP BY r.genre)
           SELECT temp2.genre, CAST(temp3.countHigh AS DOUBLE)/temp2.countAll, temp3.countHigh, temp2.countAll FROM temp2 JOIN temp3 ON temp2.genre=temp3.genre OR (temp2.genre IS NULL AND temp3.genre IS NULL)')

## find for each genre, how many artist's highest score album is their last album
## this group by is wrong... I was assuming one artist belongs to only one genre...
## actually, one album can belong to more than one genre
dbGetQuery(mydb,'WITH maxScore AS (SELECT MAX(r.score) AS maxS, r.artist,r.genre, orderIndex FROM reviewWithGenres r GROUP BY r.artist), maxOrder AS (SELECT r.score, r.artist,r.genre, MAX(orderIndex) AS maxO FROM reviewWithGenres r GROUP BY r.artist)
           SELECT maxOrder.genre, count(*) FROM maxScore JOIN maxOrder ON maxScore.artist = maxOrder.artist WHERE maxScore.OrderIndex = maxOrder.maxO GROUP BY maxOrder.genre')

dbGetQuery(mydb,'SELECT r.artist, count(DISTINCT r.genre) FROM reviewWithGenres r GROUP BY r.artist HAVING count(DISTINCT r.genre) < 3 ORDER BY count(DISTINCT r.genre) DESC LIMIT 20')

dbGetQuery(mydb,'SELECT r.genre, count(*) FROM reviewWithGenres r GROUP BY r.genre')

dbGetQuery(mydb,'SELECT r.genre, CAST(count(CASE WHEN r.score>7 THEN 1 END) AS DOUBLE) /count(*) FROM reviewWithGenres r GROUP BY r.genre')

dbGetQuery(mydb,'SELECT r.genre, count(*) FROM reviewWithGenres r WHERE r.score >7 GROUP BY r.genre')

dbGetQuery(mydb,'SELECT g.genre, count(CASE WHEN r.score>7 THEN 1 END) FROM reviews r LEFT JOIN genres g  ON r.reviewid=g.reviewid GROUP BY g.genre')

dbGetQuery(mydb,'SELECT g.genre, CAST(count(CASE WHEN r.score >7 THEN 1 END) AS DOUBLE)/count(*) AS ratio ,count(CASE WHEN r.score >7 THEN 1 END) AS count_high_score, count(*) AS count_all FROM reviews r  JOIN genres g  ON r.reviewid=g.reviewid GROUP BY g.genre')


dbGetQuery(mydb,'SELECT r.artist, r.title, r.genre, r.score FROM reviewWithGenres r WHERE r.artist = "50 cent"')


## find how many distinct artist in the table
dbGetQuery(mydb,'SELECT count(*) FROM (SELECT DISTINCT r.artist FROM reviewWithGenres r) temp')
## find how many artist that their album with highest score is their last album

dbGetQuery(mydb,'SELECT g.reviewid, g.artist, g.title FROM reviewWithGenres g WHERE g.artist IN ("muse","radiohead")')

dbGetQuery(mydb,'SELECT r.genre, count(*) AS countH FROM reviewWithGenres r GROUP BY r.genre')

## see what category is radiohead
reviewWithGenres[which(reviewWithGenres[,"artist"]=="radiohead"),]

## see what's global
head(reviewWithGenres[which(reviewWithGenres[,"genre"]=="global"),])

## find the band with at least 2 genres.
dbGetQuery(mydb,'SELECT r.artist, count(DISTINCT r.genre) FROM reviewWithGenres r GROUP BY r.artist HAVING count(DISTINCT r.genre) >=2 LIMIT 20')



dbSendQuery(mydb, 'DROP TABLE ')

```






It is not supprising that the highest score for each artist is mostly the first two album.

Be Careful on LEFT JOIN:

